---
title: 'Lesson 2: Mapping State Data'
author: "Martin Monkman"
date: "May 21, 2016"
output: html_document
---

# Ari Lamstein's U.S. Census Mapping in R Course

## Lesson 2: Mapping State Data

Install the packages, and load the data `df_pop_state` into the workspace.

```{r}

# install and load the packages
install.packages(c("choroplethr", "choroplethrMaps"))
library(choroplethr)
 
data(df_pop_state)
head(df_pop_state)

```

## Map 1 - State Population

The first map is a choropleth of state populations.

```{r cars}

state_choropleth(df_pop_state)

```

There are other options, viewed through `?state_choropleth`. More on that later.

## Map 2 - Exercise - Per-Capita Income by State

Load the more comprehensive dataset `df_state_demographics`, and assign the per capita income values to the variable `value`.

```{r}


data(df_state_demographics)
colnames(df_state_demographics)

df_state_demographics$value = df_state_demographics$per_capita_income

state_choropleth(df_state_demographics)

```

## Map 3 - Homework - Median Rent by State


```{r}

df_state_demographics$value = df_state_demographics$median_rent

state_choropleth(df_state_demographics)


```


### Map 3a (add title etc)

Add title, legend caption, constrain to 6 colors

```{r}

state_choropleth(df_state_demographics, title = "Median Rent by State",
                 legend = "Median Rent ($)", num_colors = 6)


```


### Map 3a, zoomed into west coast states 

Seven states, seven colors--so each one gets its own color in the legend.

```{r}

state_choropleth(df_state_demographics, title = "Median Rent by State: Western Region",
                 legend = "Median Rent ($)", num_colors = 7,
                 zoom = c("washington", "oregon", "california",
                          "idaho", "nevada", "utah", "arizona"))


```

