---
title: 'Lesson 4: Mapping ZIP Codes'
author: "Martin Monkman"
date: "May 25, 2016"
output: html_document
---

# Ari Lamstein's U.S. Census Mapping in R Course

## Lesson 4: Mapping ZIP Codes

Install the packages, and load the data `df_pop_state` into the workspace.

```{r message = FALSE}

setwd("C:/@data/GitHub/LamsteinCensusMapping")

# install and load the packages
install.packages("devtools")
library(devtools)

install_github("arilamstein/choroplethrZip@v1.5.0")
library(choroplethrZip)
 
data(df_pop_zip)
head(df_pop_zip)

```

## Map 1 - Population of New York State by ZIP

The first map is a choropleth of the state population by ZIP code.

```{r}

zip_choropleth(df_pop_zip, state_zoom="new york") 


```

## Map 2 - Population of Manhattan by ZIP

The first map is a choropleth of state populations. The second variant adds a reference map upderneath.

Hey look, nobody lives in Central Park.

```{r}


zip_choropleth(df_pop_zip, county_zoom=36061)


zip_choropleth(df_pop_zip, county_zoom=36061, reference_map=TRUE)

```

There are other options, viewed through `?state_choropleth`. More on that later.

## Map 2 - Map the per-capita income of each ZCTA in Manhattan

Load the more comprehensive dataset `df_zip_demographics`, and assign the per capita income values to the variable `value`.

```{r}

data(df_zip_demographics)
colnames(df_zip_demographics)


df_zip_demographics$value = df_zip_demographics$per_capita_income 
zip_choropleth(df_zip_demographics, county_zoom=36061)

```

## Homework #1 & 2


Homework #1: Use google to find the FIPS code of Los Angeles County.

According to [this source](http://county-codes.findthedata.com/l/203/Los-Angeles-County-CA), it's 06037.

Homework #2: Use df_zip_demographics and zip_choropleth to map some demographic statistic of the ZCTAs in Los Angeles county. Tweet the resulting image to me (@AriLamstein) with the hashtag #censuscourse.

Homework #3: Combine your choropleth map with a google map. Articulate one pro and one con with adding the reference map.


```{r}

df_zip_demographics$value = df_zip_demographics$median_rent 

zip_choropleth(df_zip_demographics, county_zoom=06037)


```


### add title etc

Add title, legend caption

```{r}

zip_choropleth(df_zip_demographics, county_zoom=06037,
               title = "Median Rent by ZIP Code\nLos Angeles County",
               legend = "Median Rent ($)")


zip_choropleth(df_zip_demographics, county_zoom=06037,
               title = "Median Rent by ZIP Code\nLos Angeles County",
               legend = "Median Rent ($)",
               reference_map = TRUE)



```

Pros and cons of reference maps:

* Pro - Can see the geography and its relationship with values (e.g. Catalina Island, the coast [and in particular the valuable properties in Malibu])

* Con - Underlying image can obscure the differences between ZIP code boundaries
