---
title: 'Lesson 3: Mapping County Data'
author: "Martin Monkman"
date: "May 22, 2016"
output: html_document
---

# Ari Lamstein's U.S. Census Mapping in R Course

## Lesson 3: Mapping County Data

Install the packages, and load the data `df_pop_state` into the workspace.

```{r message = FALSE}

setwd("C:/@data/GitHub/LamsteinCensusMapping")

# install and load the packages
# install.packages(c("choroplethr", "choroplethrMaps"))
library(choroplethr)
 
data(df_pop_county)
head(df_pop_county)

```

## Map 1 - State Population

The first map is a choropleth of state populations.

```{r}

county_choropleth(df_pop_county)

```

There are other options, viewed through `?state_choropleth`. More on that later.

## Map 2 - Exercise - Per-Capita Income by County

Load the more comprehensive dataset `df_state_demographics`, and assign the per capita income values to the variable `value`.

```{r}

data(df_county_demographics)

colnames(df_county_demographics)

df_county_demographics$value = df_county_demographics$per_capita_income

county_choropleth(df_county_demographics)

```

## Map 3 - Homework - Median Rent by County


```{r}

df_county_demographics$value = df_county_demographics$median_rent

county_choropleth(df_county_demographics)


```


### Map 3a (add title etc)

Add title, legend caption, constrain to 6 colors

```{r}

county_choropleth(df_county_demographics, title = "Median Rent by County",
                 legend = "Median Rent ($)", num_colors = 6)


```


### Map 3a, zoomed into Idaho 

In lesson #2 we saw that Idaho had the lowest median rent of the seven western states that were plotted. Here's a chance to take a closer look:

```{r}

county_choropleth(df_county_demographics, title = "Median Rent by County: Idaho",
                 legend = "Median Rent ($)", num_colors = 6,
                 state_zoom = c("idaho"))


```


### Map 3a.02, add a reference map to Idaho zoom 

_Revision_

After tweeting the above, Ari Lamstein tweeted back ""

So after a quick google search, I found [Combining Choropleth Maps and Reference Maps in R](http://www.arilamstein.com/blog/2015/09/30/combining-choropleth-maps-and-reference-maps-in-r/) on Ari's website.

By simply adding the parameter `reference_map = TRUE`, the choropleth now has a reference map underlying it. Easy.

```{r}

county_choropleth(df_county_demographics, title = "Median Rent by County: Idaho",
                 legend = "Median Rent ($)", num_colors = 6,
                 state_zoom = c("idaho"),
                 reference_map = TRUE)


```

### Map 3b, zoomed into Washington 

And a map of Washington.

```{r}

county_choropleth(df_county_demographics, title = "Median Rent by County: Washington",
                 legend = "Median Rent ($)", num_colors = 6,
                 state_zoom = c("washington"))


```

No surprise: the densely populated Seattle-Tacoma area, including the western side of Puget Sound, have the highest median rents in the state.
