---
title: 'Lesson 5: Mapping Historic Data'
author: "Martin Monkman"
date: "May 28, 2016"
output: html_document
---

# Ari Lamstein's U.S. Census Mapping in R Course

## Lesson 5: Mapping Historic Data

Enter the Census Bureau API key.

```{r message = FALSE}

setwd("C:/@data/GitHub/LamsteinCensusMapping")

# install and load the packages
# install.packages("devtools")

library(acs)
api.key.install("<numberhere>", file = "key.rda") 


```

## Map 1 - State Populations, 2010

The first map is a choropleth of the state population in 2010.

```{r}

library(choroplethr)

df_2010_demographics = get_state_demographics(2010)
df_2010_demographics$value = df_2010_demographics$total_population

state_choropleth(df_2010_demographics, 
                 title  = "2010 State Population Estimates",
                 legend = "Population")

```

## Map 2 - County Populations, Texas, 2011


```{r}

df_2011_demographics = get_county_demographics(2011)
df_2011_demographics$value = df_2011_demographics$percent_hispanic

county_choropleth(df_2011_demographics,
                  title      = "Texas 2011 County\nPercent Hispanic",
                  legend     = "Percent Hispanic",
                  state_zoom = "texas")


```


## Homework - Mapping San Francisco demographics by ZIP code, 2011

Mapping median rents in the San Francisco County, using the 2011 ACS.

The final version of the map adds a reference map, and divides the rents into quintiles.


```{r}

library(choroplethrZip)

df_zip_demographics2011 <- get_zip_demographics(endyear = 2011, span = 5)
colnames(df_zip_demographics2011)


df_zip_demographics2011$value = df_zip_demographics2011$median_rent
colnames(df_zip_demographics2011)

zip_choropleth(df_zip_demographics2011, county_zoom=6075)


zip_choropleth(df_zip_demographics2011, county_zoom=6075,
               title = "Median Rent by ZIP Code, 2011\nSan Francisco County",
               legend = "Median Rent ($)",
               num_colors = 5,
               reference_map = TRUE)


```




## Homework #1 & 2


Homework #1: Use google to find the FIPS code of Los Angeles County.

According to [this source](http://county-codes.findthedata.com/l/203/Los-Angeles-County-CA), it's 06037.

Homework #2: Use df_zip_demographics and zip_choropleth to map some demographic statistic of the ZCTAs in Los Angeles county. Tweet the resulting image to me (@AriLamstein) with the hashtag #censuscourse.

Homework #3: Combine your choropleth map with a google map. Articulate one pro and one con with adding the reference map.


```{r}

df_zip_demographics$value = df_zip_demographics$median_rent 

zip_choropleth(df_zip_demographics, county_zoom=06037)


```


### add title etc

Add title, legend caption

```{r}

zip_choropleth(df_zip_demographics, county_zoom=06037,
               title = "Median Rent by ZIP Code\nLos Angeles County",
               legend = "Median Rent ($)")


zip_choropleth(df_zip_demographics, county_zoom=06037,
               title = "Median Rent by ZIP Code\nLos Angeles County",
               legend = "Median Rent ($)",
               reference_map = TRUE)



```

Pros and cons of reference maps:

* Pro - Can see the geography and its relationship with values (e.g. Catalina Island, the coast [and in particular the valuable properties in Malibu])

* Con - Underlying image can obscure the differences between ZIP code boundaries
